Puck Predictions: Unraveling the NHL Game Forecasting Riddle

### 2024-04-12 Latex/draft.tex
Pushing The Good Changes(TM)

The models were each trained on prior seasons, and tested on the most recent 2022-2023 season. This split was chosen because in practice our model would have access to historical data but not have access to the current season's data as it happens.

@@

\begin{figure}
    \centering
    \includegraphics[width=0.9\textwidth]{images/good_bayesian_example.png}
    \caption{Win probability graph using the Bayesian Network. As seen above, the win probability stayed rather stagnant until a goal was scored, where the probability shifted.}
    \label{fig:bayesian_graph}
\end{figure}

@@

\begin{figure}
    \centering
    \subfigure[]{\includegraphics[width=0.6\textwidth]{images/good_xgb_example.png}} 
    \subfigure[]{\includegraphics[width=0.6\textwidth]{images/good_xgb_example(2).png}} 
    \caption{For both of these plots, the home team is above the the 50\% line, with the
    away team below. The title has the final score of the game. \\
        \textbf{(a)} The live win probabilities for the Nashville Predators at the Vegas Golden Knights. 
    Notice the probability shifts after one team scores a goal.  \\
        \textbf{(b)} The live win probabilities for the Calgary Flames at the Winnipeg Jets.
        Notice the stark changes in probabilities after each goal. Also notice the
        shift in probabilities even when no goals are scored.}
    \label{fig:xgboost_res}
\end{figure}

@@

In Table~\ref{tab:game_accuracies} we note that each model increased in accuracy as the game progressed. 
The Bayesian network and MCMC model were the most accurate, while the XGBoost model had lower accuracy.
We infer that the XGBoost models were overfit, as on the training data it would regularly achieve higher than 90\% accuracy at most points during the game. 
Also, during experimentation with the models, we noticed that the MCMC simulator tended to produce higher probabilities for the away team winning, 
and the XGBoost model tended to give higher probabilities to the home team winning. 
The Bayesian Network was unbiased towards the home or away team. 
We trained the MCMC model on the flow of events in the game, so it appears that the model picked up on a more predictive flow in how away teams play. 
Though the MCMC model was generally the most accurate, it takes about 2 minutes to simulate 50 games and get a single prediction while XGBoost and the Bayesian Network can make a prediction in under two seconds.

>>>>

The models were each trained on prior seasons, and tested on the most recent 2023-2024 season. This split was chosen because in practice our model would have access to historical data but not have access to the current season's data as it happens.

@@

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{images/good_bayesian_example.png}
    \caption{Win probability graph using the Bayesian Network. As seen above, the win probability stayed rather stagnant until a goal was scored, where the probability shifted.}
    \label{fig:bayesian_graph}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{images/good_xgb_example.png}
    \caption{We can see a high variance of win probability during this game. It is interesting to
    see how a goal being scored impacts the win probability. It is also interesting to see how
    the probability shifts without a goal being scored, implying the other events in the game
    are important to the win probability.}
    \label{fig:xgboost_res}
\end{figure}

@@

In Table~\ref{tab:game_accuracies} we note that each model increased in accuracy as the game progressed. 
The Bayesian network and MCMC model were the most accurate, while the XGBoost model had lower accuracy.
We infer that the XGBoost models were overfit, as on the training data it would regularly achieve higher than 90\% accuracy at most points during the game. 
Also, during experimentation with the models, we noticed that the MCMC simulator tended to produce higher probabilities for the away team winning, 
and the XGBoost model tended to give higher probabilities to the home team winning. 
The Bayesian Network was unbiased towards the home or away team. 
We trained the MCMC model on the flow of events in the game, so it appears that the model picked up on a more predictive flow in how away teams play. 
Though the MCMC model was generally the most accurate, it takes about 2 minutes to simulate 50 games and get a single prediction while XGBoost and the Bayesian network can make a prediction in under two seconds.
